---
title: "Notes"
output: pdf_document
date: "2025-03-20"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
dict <- read.csv("final_dictionary.csv")
formula <- read.csv("us_news_formula.csv")
us_formula_weights <- read.csv("us_news_weights.csv")
predictions <- read.csv("final_predictions.csv")
```

```{r}
# graph to compare weights of each category from our and us news formula
weights_by_cat <- dict %>%
  group_by(category) %>%
  summarize(weights = sum(weights_rounded)) %>%
  print()

weights_by_cat$weights[8] <- 0.12
weights_by_cat <- weights_by_cat[-1,]
weights_by_cat <- weights_by_cat %>%
  mutate(From = "ours")

us_formula_weight <- us_formula_weights[-7,]
us_formula_weight <- rbind(us_formula_weight, data.frame(category = "affordability", weight = 0))
us_formula_weight <- us_formula_weight %>%
  mutate(weights = weight) %>%
  select(c("weights", "category")) %>%
  mutate(From = "original")

weights_compare <- rbind(us_formula_weight, weights_by_cat)

ggplot(weights_compare, aes(fill = From, x = category, y = weights*100)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(fill = "Formula", y = "Weights (%)", x = "Category")
```

## Notes

- The borrower debt variable is sourced from college scorecard (source: https://www.usnews.com/education/best-colleges/articles/ranking-criteria-and-weights)
- US News explicitly said no admission rates, but our model had one variable for admission rate
- weird variables: no pell withdraw, forbearance, 
- there were some diversity-related stuff

- US news factors that we know are included in our model (exact):
  - **Faculty salaries**: 
  - **Full-time faculty**: 
  - **Financial resources per student**:
  - **Pell graduation rate**:
  
- Things that are not directly on the formula but kinda adjacent:
  - **Graduation rates**: 
  - **Retention rate**: 
  
- New things
  - location, what degrees they offer, cost of attendance, median debt for independent students, faculty diversity


## US News Variables

#### Comple
- **Graduation rates** (16%): percentage of entering full-time first-year bachelor's or equivalent degree-seeking students who completed a bachelor's or equivalent degree within a six-year period (averaged over four years)
- **First-year retention rates** (5%): average proportion of the first-year classes entering who returned to school the following fall
- **Graduation rate performance** (10%): four-year rolling average comparing each college's six-year graduation rates with what U.S. News predicted for their corresponding start year through entering classes. If the actual graduation rate is higher than the predicted rate, the college is enhancing achievement or is overperforming. If its actual graduation rate is lower than the predicted rate, it's underperforming.

#### Mobility
- **Pell graduation rates** (5.5%): six-year graduation rates of students who received Pell Grants
- **Pell graduation performance** (5.5%): success at achieving equitable outcomes for students from underserved backgrounds. Initial score is calculated by comparing Pell graduation rates to non-Pell graduating rates (highest is 1), and schools whose entering cohorts were at least 50% Pell Grant students had their scores augmented by 0.5; schools below 50% Pell Grant students had their scores augmented by the proportion that received Pell Grants.
- **College grads earning more than a high school grad** (5%): proportion of a school's federal loan recipients who in five years since completing their undergraduate degrees were earning more than the median salary of a 25-34yo whose highest level of education is high school

#### Debt
- **Borrower debt** (5%): each school's typical average accumulated federal loan debt among only borrowers who graduated

#### Faculty
- **Faculty salaries** (6%)(+2%): averaged salaries from all full-time instructional tenured and nontenured faculty
- **Full-time faculty** (2%)(+1%): proportion of the fall instructional faculty that was full time
- **Student-faculty ratio** (3%)(+1%): ratio of full-time-equivalent students to full-time-equivalent faculty members during fall, excluding faculty that only teaches graduate level

**Others**
- **Standardized tests**(5%): assessed schools on their fall 2023 SAT/ACT scores if they were reported on at least half of their new entrants, if less than half this category is cancelled and weights of graduation rates is increase by 5%
- **Peer assessment** (20%)

**Academic**
- **Financial resources per student** (8%): represents a school's ability to have a strong environment for instruction and impact in academia, measured by comparing an institution's total expenditures on instruction, research, public service, academic support, student services and institutional support against its total academic year full-time equivalent student enrollment (only academic spending).


